---
title: "Project X: Data Analysis"
shorttitle: "Short Project"
project: "Project"
srpt_number: 007
version: "`r CosmosReportTemplates::git_tag()`"
author: ''
date: "`r format(Sys.time(), '%d %B, %Y')`"
investigators:
  - name: ""
  - name: ""
project_team:
analysis_team:
analysis_contact:
organization: "Analysis Group"
output:
  word_document: default
  CosmosReportTemplates::bbsr_article_pdf: default
  pdf_document: default
  html_document: default
toc: yes
---


# Overview

The goal of this analysis is to examine the correlation of individual probesets to the RSI score, across several large datasets. We will use three different datasets: TCC (HuRSTA array), EXPO (U133 arrays) and CCLE (U133 arrays).

```{r setup}
suppressPackageStartupMessages({
  library(readxl)
  library(knitr)
  library(kableExtra)
  library(ggplot2)
  library(dplyr)
  library(Biobase)
})
```

# Load Data
Load the three different datasets.
```{r load_data}
ccle<-readRDS(file="../data.raw/ccle_cel.rds")
expo<-readRDS(file="../data.raw/expo_rma_exprs.rds")
tcc<-readRDS(file="../data.raw/tcc.primary.rds")

translation<-read_excel("../data.raw/Translation.xlsx")
```

# Run correlations
```{r}

ccle_correlations<-apply(exprs(ccle), 1, function(y){cor(y, ccle$RSI)})
expo_correlations<-apply(exprs(expo), 1, function(y){cor(y, expo$rsi)})
tcc_correlations<-apply(exprs(tcc), 1, function(y){cor(y, tcc$RSI)})
```

# Analysis of correlations

This shows that most genes are not correlated to RSI.
```{r}
hist(ccle_correlations, main="CCLE")
hist(expo_correlations, main="EXPO")
hist(tcc_correlations, main="TCC")
```

What probesets are correlated at a modest ```R=0.5```?
```{r}
kable(tcc_correlations[which(abs(tcc_correlations)>0.5)], col.names=c("R"), caption="TCC Correlated probesets", booktabs=TRUE)


kable(expo_correlations[which(abs(expo_correlations)>0.5)], col.names=c("R"), caption="EXPO correlated probesets", booktabs=TRUE)

kable(ccle_correlations[which(abs(ccle_correlations)>0.5)], col.names=c("R"), caption="CCLE correlated probesets", booktabs=TRUE)

```


```{r}
	u133_coeffs<-c("AFFX-HUMISGF3A/M97935_MA_at"=0.0254552046046931,
			"201209_at" = -0.020446868981449,
			"201466_s_at" = 0.0128282697929755,
			"201783_s_at" = -0.00381713588908642,
			"202123_s_at" = 0.107021274358643,
			"202531_at" = -0.0441682719957224,
			"205962_at" = -0.00924314364533049,
			"207957_s_at" = -0.00175888592095915,
			"208762_at" = -0.00025093287356907,
			"211110_s_at" = -0.00980092741843779)

  hursta_coeffs<-c("merck-NM_139266_at"=0.0254552046046931,
			"merck-NM_004964_at" = -0.020446868981449,
			"merck-NM_002228_at" = 0.0128282697929755,
			"merck2-BC069248_at" = -0.00381713588908642,
			"merck-NM_007313_s_at" = 0.107021274358643,
			"merck-NM_002198_at" = -0.0441682719957224,
			"merck-BQ646444_a_at" = -0.00924314364533049,
			"merck2-X06318_at" = -0.00175888592095915,
			"merck-NM_001005782_s_at" = -0.00025093287356907,
			"merck-NM_000044_a_at" = -0.00980092741843779)
```


# Gene Correlations
Show the correlations of the ten genes.

## TCC
```{r}
for (g in names(hursta_coeffs)) {
  myplot<-ggplot(data.frame(gene=exprs(tcc)[g,], RSI=tcc$RSI, tissue=tcc$SOO_Conformed), aes(x=gene, y=RSI, color=tissue)) +
    geom_point() +
    theme_bw() +
    ggtitle(g)
  
  print(myplot)
}
```

## CCLE
```{r}
for (g in names(u133_coeffs)) {
  myplot<-ggplot(data.frame(gene=exprs(ccle)[g,], RSI=ccle$RSI, tissue=ccle$`Site Primary`), aes(x=gene, y=RSI, color=tissue)) +
    geom_point() +
    theme_bw() +
    ggtitle(g)
  
  print(myplot)
}
```

## EXPO
```{r}
for (g in names(u133_coeffs)) {
  myplot<-ggplot(data.frame(gene=exprs(expo)[g,], RSI=expo$rsi, tissue=expo$source_name_ch1), aes(x=gene, y=RSI)) +
    geom_point() +
    theme_bw() +
    ggtitle(g)
  
  print(myplot)
}
```

# ABL-IRF1

# TCC
```{r}
ABL1<-translation %>% filter(`Gene Symbol`=="ABL1") %>% pull("HuRSTA Array")
IRF1<-translation %>% filter(`Gene Symbol`=="IRF1") %>% pull("HuRSTA Array")

ratio<-exprs(tcc)[ABL1,]-exprs(tcc)[IRF1,]

ggplot(data.frame(ratio=ratio, rsi=tcc$RSI, tissue=tcc$SOO_Conformed), aes(x=ratio, y=rsi, col=tissue)) +
  geom_point() +
  theme_bw() +
  annotate(geom="text",x=-5,y=0.75, label=sprintf("R=%5.2f",cor(ratio, tcc$RSI)))

  
```

# CCLE
```{r}
ABL1<-translation %>% filter(`Gene Symbol`=="ABL1") %>% pull("HG-U133 Plus 2.0 Probeset")
IRF1<-translation %>% filter(`Gene Symbol`=="IRF1") %>% pull("HG-U133 Plus 2.0 Probeset")

ratio<-exprs(ccle)[ABL1,]-exprs(ccle)[IRF1,]

ggplot(data.frame(ratio=ratio, rsi=ccle$RSI, tissue=ccle$`Site Primary`), aes(x=ratio, y=rsi, col=tissue)) +
  geom_point() +
  theme_bw() +
  annotate(geom="text",x=-5,y=0.75, label=sprintf("R=%5.2f",cor(ratio, ccle$RSI)))

  
```

# expo
```{r}
ABL1<-translation %>% filter(`Gene Symbol`=="ABL1") %>% pull("HG-U133 Plus 2.0 Probeset")
IRF1<-translation %>% filter(`Gene Symbol`=="IRF1") %>% pull("HG-U133 Plus 2.0 Probeset")

ratio<-exprs(expo)[ABL1,]-exprs(expo)[IRF1,]

ggplot(data.frame(ratio=ratio, rsi=expo$rsi, tissue=expo$source_name_ch1), aes(x=ratio, y=rsi)) +
  geom_point() +
  theme_bw() +
  annotate(geom="text",x=-5,y=0.75, label=sprintf("R=%5.2f",cor(ratio, expo$rsi)))

  
```


```{r session.info.table,results='asis',echo=FALSE}
CosmosReportTemplates::print_session_information()
```
